Experience Report: Haskell in Computational Biology

### 2012-07-01 mrfy_experience_report.tex
sensible pdf build without latexmk
A couple remarks.

Our code represents a query sequence as an immutable vector of amino
acids.
In~idiomatic Haskell, 
we might represent an individual amino acid~$x_i$
using a value of algebraic data type:
\begin{smallverbatim}
data Amino = Ala | Cys | Asp | Glu | Gly | ...
\end{smallverbatim}
But our models use a
legacy file format that represents each amino acid as a small integer.
Our representation is therefore
\smallverbatiminput{aa}
The log of a probability is a negative number.
To~make these numbers more pleasant to read, we negate them,
which eliminates hordes of minus signs from our file formats.
The negated logarithm of a probability is called a \emph{score}.
\smallverbatiminput{score}
Type \texttt{Score} has a limited \texttt{Num} instance which permits
scores to be added and subtracted but not multiplied.

@@

Ideally, MRFy would use now-classic, lazy, modular
technique advocated by \citet{hughes:why}, in which one function
computes an infinite sequence of points, and another function
uses a finite prefix to decide on an approximation.
But because MRFy's search is stochastic, 
making MRFy's search lazy and modular is not so easy.

>>>>

Our code represents a query sequence as an immutable vector of amino
acids.
In~idiomatic Haskell, 
we might represent an individual amino acid~$x_i$
using a value of algebraic data type:
\begin{smallverbatim}
data Amino = Ala | Cys | Asp | Glu | Gly | ...
\end{smallverbatim}
But our models use a
legacy file format that represents each amino acid as a small integer.
Our representation is therefore
\smallverbatiminput{aa}

\remark{I'm not sure if this is the whole story, since the reader might simply
  ask, ``Why not map 0 to Ala, 1 to Cys, etc.?''. The real reason we keep the
  ``Int'' here I think, is that it provides easy indexing into the match and
  insertion emission vectors.}

The log of a probability is a negative number.
To~make these numbers more pleasant to read, we negate them,
which eliminates hordes of minus signs from our file formats.
The negated logarithm of a probability is called a \emph{score}.
\smallverbatiminput{score}
Type \texttt{Score} has a limited \texttt{Num} instance which permits
scores to be added and subtracted but not multiplied.

@@

\remark{The opening sentence here sounds awkward, but I'm not sure if it's
  because I'm reading it wrong. It seems like ``technique'' should be
  pluralized, or perhaps ``would use now-classic: a lazy and modular
  techniqe''.}
Ideally, MRFy would use now-classic, lazy, modular
technique advocated by \citet{hughes:why}, in which one function
computes an infinite sequence of points, and another function
uses a finite prefix to decide on an approximation.
But because MRFy's search is stochastic, 
making MRFy's search lazy and modular is not so easy.
